# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Collaborative Filtering Engine

Author: Daniel Lopez

========================================================

# Initialization

Import libraries

```{r}
# Used for movie ratings visualization
library(Amelia)
# Exercises 1 to 5 should not need new libraries
# See functions: 
# - sort
# - apply
# - colMeans
# - subset
# Used for Movielens analysis
library(dplyr)
library(ggplot2)
library(tidyr)
library(gridExtra)
library(corrplot)
library(magrittr)
library(tidyverse)


```

Load the critics dataset.

```{r}

critics = read.csv("C:/Users/Daniel/Desktop/Recommendation Engines/Lab 1 Deliverable/critics.csv")


```

# 1) Pearson correlation coefficient 
Calculate the Pearson correlation coefficient between Victoria and Nuria

**Expected correlation between Victoria and Nuria**: 0.3651484

```{r fig.width=7, fig.height=6}

vic <- critics[15,2:21]

nur <- critics[12,2:21]

cor.test(t(nur),t(vic))

```

# 2) Compare two movie critics
Using the calculated Pearson correlation in the previous section. Compare and show the movie ratings for Victoria and Nuria on a scatter plot. How is this compared with the same ratings of Maria and Nerea on a scatter plot? What about Chris and Jim? What it means?

```{r fig.width=7, fig.height=6}

# Plot test for Nuria and Victoria



plot(t(nur),t(vic),xlab= "Rating's for Nuria",ylab = "Rating's for Victoria",
     main="Rating comparison between Users",col= (c("red","green")),pch= 16)

# Plot for Maria and Nerea

plot(t(critics[2,2:21]),t(critics[13,2:21]),xlab= "Rating's for Maria",ylab = "Rating's for Nerea",
     main="Rating comparison between Users",col= (c("orange","purple")),pch= 16)


# Plot for Chris and Jim

plot(t(critics[10,2:21]),t(critics[9,2:21]),xlab= "Rating's for Chris",ylab = "Rating's for Jim",
     main="Rating comparison between Users",col= (c("blue","black")),pch= 16)



```

# 3) Top Recommendations 
Get the top 5 movies recommended for Victoria. by using a weighted average of every other critic's rankings.

**Expected result**:
3.7917013044215, 'The Matrix'
3.50776533175371, 'Forrest Gump'     
3.33118834864677, 'The Sixth Sense' 
3.11491825315719, 'Shakespeare in Love'
2.9124513228665, 'Blade Runner'

```{r fig.width=7, fig.height=6}

# First we need to group the data

grouped_critics <- gather(critics,"film","rating",-User)

grouped_critics$rating <- as.numeric(grouped_critics$rating)

# Now we need to assign the average rating per user and movie 

grouped_critics <- grouped_critics %>%
  group_by(User) %>%
  mutate(avg_rating = mean(rating, na.rm = TRUE)) %>%
  ungroup()

# We get the ratings that Victoria gave

Victoria_ratings <- grouped_critics %>%
  filter(User == "Victoria") %>%
  extract2("rating")

# We include the correlation

grouped_critics <- grouped_critics %>%
  group_by(User) %>%
  mutate(correlation = cor(rating, Victoria_ratings,use="complete.obs")) %>%
  ungroup()

# We get the info just for Vic

victoria_movies <- subset(grouped_critics, User == "Victoria")

victoria_movies <-victoria_movies[is.na(victoria_movies$rating),]

films_not_seen <- victoria_movies[1:7,2]

# Possible recommendations for Victoria
# Films not seen with raitings
df <- grouped_critics[grouped_critics$film %in% films_not_seen$film,]

# We calculate the recomm_score for the total movies

df <- df %>%
  group_by(film) %>%
  summarise(recomm_score = sum((rating - avg_rating) * correlation, na.rm = TRUE) / (sum(abs(correlation)) - 1)) %>%
    mutate(recomm_score = recomm_score + mean(Victoria_ratings, na.rm = TRUE))

# Now we order from high to low

df <- df %>% arrange(desc(recomm_score))

head(df,5)

```

# 4) Top similar critics
Return the top 5 critics that are most similar to Victoria. 

**Expected result**:
0.9449112, 'Rachel'
0.5976143, 'Ana' 
0.5789794, 'Oriol'
0.4925922, 'Maria'
0.4273247, 'Carles'

```{r fig.width=7, fig.height=6}

# We get the movies watched by victoria

victoria_movies_s <- subset(grouped_critics, User == "Victoria")

victoria_movies_s <-victoria_movies_s[!is.na(victoria_movies_s$rating),]

films_seen <- victoria_movies_s[1:13,2]

# Dataframe with those that watched the same movies as Victoria

df_user <- grouped_critics[grouped_critics$film %in% films_seen$film,]

# We take Victoria out of the set

df_user <- df_user %>% filter( User != "Victoria")

# I only want those users who have seen the same movies as VIc

df_user <- df_user[!is.na(df_user$rating),]

# We only need the name and cor columns

df_user <- df_user[,c(1,5)]

# We remove the duplicate names

df_user <- unique(df_user[,])

# Now we order from high to low

df_user <- df_user %>% 
            arrange(desc(correlation))

head(df_user,5)

```